---
title: "R Notebook"

---

# import libraries

```{r}
library(here)
library(dplyr)
library(tidyr)
library(readxl)
library(lubridate)
# library(ggplot2)
```

# read 5th-wave data

```{r}
read_xlsx_from_folder = function(filepath, match_by = NULL){
  filenames = list.files(filepath)
  # browser()
  filenames = filenames[grepl(match_by, filenames)]
  df = as.data.frame(c())
  for(name in filenames){
    xlsx = readxl::read_xlsx(paste0(filepath, name), col_types = "text")
    df = bind_rows(df, xlsx)
  }
  df
}
```

# COVID drug use 
# 25 May 2022: jan data is missing and we dont care, becos antiviral was dispensed only after 25th feb

```{r}
# directory as in onelook drive: ~/COVID Database/Patient-based Record/Antiviral (List of patients & drug use)/
# also antimicrobial folder

antiviral = read_xlsx_from_folder("./data/5th Wave 202202/Antiviral/", ".xlsx")
antiviral = antiviral %>%
  # filter by only paxlovid and molnupiravir
  filter(grepl("PAXLOVID|MOLNUPIRAVIR", `Drug Name`))
# steroid administration route data
# directory as in onelook drive: ~/COVID Database/Patient-based Record/XLSX (Preferred)/Wave 5/Systemic Corticosteroid/
steroid = read_xlsx_from_folder("./data/Wave 5/Steroid/", "_DRG.xlsx")
# baricitinib 
# directory as in onelook drive: ~/COVID Database/Patient-based Record/XLSX (Preferred)/Wave 5/Interferons & Antibodies/
antibodies = read_xlsx_from_folder("./data/Wave 5/Antibodies/", "_DRG.xlsx")

```

# meds status
# onelook drive: ~/COVID Database/Patient-based Record/XLSX (Preferred)/Wave 5/Premorbid/Drugs

```{r}
# need filter by drug names
antiplatelet = read_xlsx_from_folder("./data/Drugs/CVS - Antiplatelet  & Coagulants/", "Dx")
ace_inhibitor = read_xlsx_from_folder("./data/Drugs/CVS - Renin-Angiotensin/", "Dx")
beta_blocker = read_xlsx_from_folder("./data/Drugs/CVS - Beta-blocker/", "Dx")
calcium_channel_blockers = read_xlsx_from_folder("./data/Drugs/CVS - Calcium-channel Blocker (tbc)/", "Dx")
diuretics = read_xlsx_from_folder("./data/Drugs/CVS - Diuretics/", "Dx")
statins = read_xlsx_from_folder("./data/Drugs/CVS - Lipid (tbc)/", "Dx")
system_corticosteroid = read_xlsx_from_folder("./data/Drugs/Endocrine - Corticosteroids/", "Dx")
antidiabetics = read_xlsx_from_folder("./data/Drugs/Endocrine - Diabetes/", "Dx")
rheumatoids = read_xlsx_from_folder("./data/Drugs/Rheumatology/", "Dx")

# no need filter by drug names
broncodilators = read_xlsx_from_folder("./data/Drugs/Respiratory - bronchodilators/", "Dx")
cromoglycate = read_xlsx_from_folder("./data/Drugs/Respiratory - Cromoglycate/", "Dx")
corticosteroid = read_xlsx_from_folder("./data/Drugs/Respiratory - Inhaled Corticosteroid/", "Dx")

broncodilators = bind_rows(broncodilators, cromoglycate, corticosteroid)

cancer = read_xlsx_from_folder("./data/Drugs/Malignancy/", "Dx")

```

# COVID patients IP data, which contains sex, age, icd9 diagnosis, drug use, death and length of stay data

```{r}
options(warn = -1)
# directory as in onelook drive: ~/COVID Database/Patient-based Record/XLSX (Preferred)/Wave 5/
ip1 = read_xlsx_from_folder("./data/Wave 5/A&E and Hospitalization/202201/", "_IP.xlsx")
ip2 = read_xlsx_from_folder("./data/Wave 5/A&E and Hospitalization/202202/Dx/", "_IP.xlsx")
ip3 = read_xlsx_from_folder("./data/Wave 5/A&E and Hospitalization/202203/Dx/", "_IP.xlsx")
ip4 = read_xlsx_from_folder("./data/Wave 5/A&E and Hospitalization/202204/", "_IP.xlsx")
ip5 = read_xlsx_from_folder("./data/Wave 5/A&E and Hospitalization/202205/", "_IP.xlsx")
ip6 = read_xlsx_from_folder("./data/Wave 5/A&E and Hospitalization/202206/", "_IP.xlsx")
ip7 = read_xlsx_from_folder("./data/Wave 5/A&E and Hospitalization/202207/", "_IP.xlsx")
ip = bind_rows(ip1, ip2, ip3, ip4, ip5, ip6, ip7)
ip1 = ip2 = ip3 = ip4 = ip5 = ip6 = ip7 = NULL
# filter only 22nd feb - 15th jul
ip = ip %>% 
  filter(ymd(`Admission Date (yyyy-mm-dd)`) >= "2022-02-22",
         ymd(`Admission Date (yyyy-mm-dd)`) <= "2022-07-15")
```


```{r}
# directory as in onelook drive: ~/COVID Database/Patient-based Record/XLSX (Preferred)/Wave 5/
# two separate death cause folders
death = read_xlsx_from_folder("./data/Death Cause/", ".xlsx")
death = death %>%
  select(-`No. of Episodes`) %>%
  distinct()

sdi = readxl::read_excel("./SDI_map.xlsx")
sdi = sdi %>%
  mutate(sdi_q = cut(average_SDI_2, breaks = quantile(average_SDI_2), include.lowest = T))
```

# import organ dysfunction 

```{r}
# From Luke's
organ_dys = read.csv("./culture_results/paper3_organ_dysfunction_outcomes.csv")

organ_dys %>% count(type)

organ_dys = organ_dys %>% 
  mutate(min_date = as.Date(min_date),
         Reference.Key = as.character(Reference.Key)) %>%
  pivot_wider(names_from = type,
              values_from = min_date) %>%
  rename(`Reference Key` = Reference.Key,
         "Circulatory.shock" = "Circulatory shock",
         "Respiratory.failure" = "Respiratory failure",
         "Acute.kidney.injury" = "Acute kidney injury",
         "Coagulopathy"="Coagulopathy",
         "Acute.liver.impairment" = "Acute liver impairment") 
```

# import culture data

```{r}
culture = read_xlsx_from_folder("./data/Wave 5/Culture/", ".xlsx")
culture.id = culture %>%
  select(`Reference Key`, 
                    culture.date = "Bacterial Culture and ST Within Hospital Stay - Ref. Date (yyyy-mm-dd HH:MM)",
                    name = "Bacterial Culture and ST Within Hospital Stay - Organism Description (Corporate)",
         specimen = "Bacterial Culture and ST Within Hospital Stay - Specimen") %>%
  mutate(culture.date = as.Date(culture.date)) %>%
  filter(
    # blood only
    grepl("Blood",specimen),
    # remove irrelevant bacteria
         !name %in% c("Staphylococcus, coagulase negative", "Staphylococcus epidermidis", "Diphtheroids" ,
                      "Staphylococcus capitis", "Staphylococcus caprae", "Bacillus cereus group",
                      "Bacillus cereus", "Clostridium perfringens", "Clostridium ramosum", "Paenibacillus species",
                      "Bacillus species", "Gram positive bacilli")) 
  # select(-specimen, -name)

# culture.id %>% count(name) %>%
#   arrange(desc(n)) %>%
#   writexl::write_xlsx(., path = "culture.blood.positive.pp3.xlsx")

```

# icu data (not used)

```{r}
demo_ip = read_xlsx_from_folder("./data/5th Wave 202202/", "_demo_IP.xlsx")
icu = demo_ip %>%
  filter(grepl("ICU", `Clinical Specialty (EIS)`),
         grepl("Transfer In", `Movement Type`)) %>%
  select(`HN Number`, `Admission Date (yyyy-mm-dd)`, `Movement Time`) %>%
  mutate(daystillicu = date(ymd_hm(`Movement Time`)) - ymd(`Admission Date (yyyy-mm-dd)`))
icu = aggregate(.~`HN Number`, icu, min)  %>%
  select(-`Admission Date (yyyy-mm-dd)`, -`Movement Time`)
icu %>% count(`HN Number`) %>% filter(n > 1)
```

# antibiotics

```{r}
antibiotics = read_xlsx_from_folder("./data/Wave 5/Antimicrobial/", ".xlsx")
antibiotics = antibiotics %>%
  filter(`Dispensing Date (yyyy-mm-dd)`>=as.Date("2022-02-22"))
```

