---
title: "R Notebook"

---

# import libraries

```{r}
library(survival)
library(broom)
library(survey)
library(comorbidity)
library(WeightIt)
library(cobalt)
library(tableone)
library(ggfortify)
library(dplyr)
library(tidyr)
source("./scripts/helper.R")
```


# 2022-04-26: episode death info is missing from hosp files; 
# please refer to data_import.Rmd for IP data import 
# check drug names (optional)

```{r}
unique(ip$`Drug Name`)[grepl("DEXAMETHASONE", unique(ip$`Drug Name`))]
```

# dissociate drug use and ip data (unique by HN number)

```{r}
# specify icd9 code column index
diag_cols = 35:49

ip.nodrug = ip %>%
  select(-`Drug Name`) %>%
  distinct() %>%
  # exclude pregnancy
  mutate(pregnancy = ifelse(if_any(diag_cols, ~between(., "630", "679.99")), T, F),
         pregnancy = ifelse(is.na(pregnancy), F, pregnancy)) %>%
  filter(!pregnancy) %>%
  select(-pregnancy)
```

# identification dataframe of IP patients' first admissions only (ip.id)

```{r}
ip.id = ip.nodrug %>%
  select(`Reference Key`, `HN Number`, `Admission Date (yyyy-mm-dd)`, `Discharge Date (yyyy-mm-dd)`) %>%
  group_by(`Reference Key`) %>%
  slice(which.min(ymd(`Admission Date (yyyy-mm-dd)`))) %>%
  ungroup()

writexl::write_xlsx(ip_first_by_patient_id, "./ip.id.xlsx")
```

# ip.sexage

```{r}
ip.sexage = ip.nodrug %>%
  select(`Reference Key`, Sex, age = `Admission Age (Year) (episode based)`) %>%
  # remove NA in age, no NA in sex
  filter(!is.na(age)) %>%
  mutate(age = as.numeric(age),
         age_grp3 = cut(age, breaks = c(0, 65, 75, Inf)),
         age_grp5 = cut(age, breaks = c(0, 60, 65, 70, 75, 80, Inf))) %>%
  group_by(`Reference Key`) %>%
  # get min age
  slice(which.min(age)) %>%
  ungroup()

writexl::write_xlsx(ip.sexage, path = "./ip.sexage.xlsx")
```

# comorbs, change to comorbidity package (ip.diag)

```{r}
# specify icd9 code column index
diag_cols = 35:49

ip.diag = ip.nodrug %>% 
  select(`HN Number`, diag_cols) %>%
  impute_comorbs(., diag_cols - 33) 

ip.diag = aggregate(ip.diag, by = list(ip.diag[, "HN Number"]), FUN = "max") %>%
  select(-Group.1)

ip.charlson = ip.nodrug %>% 
  select(`HN Number`, diag_cols) %>%
  # filter(`HN Number` %in% ip_first_by_patient_id$`HN Number`) %>%
  pivot_longer(cols = diag_cols - 33 , names_to = "diagnosis", values_to = "code") %>%
  drop_na() %>%
  group_by(`HN Number`) %>%
  distinct() %>%
  ungroup() %>%
  mutate(id = 1:n())

# CCI
charlson = comorbidity::comorbidity(x = as.data.frame(ip.charlson), 
                                   id = "id", 
                                   code = "code", 
                                   map = "charlson_icd9_quan", 
                                   assign0 = F) 
# charlson comorbs and cci by HN Number
charlson = charlson %>%
  left_join(select(charlson_diag, `HN Number`, id))
charlson = aggregate(charlson, by = list(charlson[, "HN Number"]), FUN = "max") %>%
  select(-Group.1)
charlson = charlson %>% 
  rowwise() %>% 
  mutate(cci = sum(c_across(mi:aids))) %>%
  mutate(cci = case_when(
    cci == 0~ "0",
    cci < 3~ "1-2",
    cci < 5~ "3-4",
    cci >= 5~ "5+",
    TRUE ~ "0"
  ))
cci = charlson %>%
  select(`HN Number`, cci)
# writexl::write_xlsx(charlson_diag, path = "./charlson_diag.xlsx")


ip.diag = ip.diag %>%
  left_join(cci, by = "HN Number") %>%
  # replace_na of ci
  mutate(cci = ifelse(is.na(cci), "0", cci))
  # select(-diagnosis, -id, -code)
  
unique(ip_diag$cci)

# check if there is any NA, missing data
ip.diag %>% filter_all((any_vars(is.na(.))))  
writexl::write_xlsx(ip.diag, "./ip.diag.xlsx")
```

# 28d mortality and length of stay
# respiratory mortality: ICD-10 code J00-J99, and U071

```{r}

# impute 28d death
ip.death = ip.id %>%
  left_join(select(ip.nodrug, `HN Number`, `Admission Date (yyyy-mm-dd)`,
                   `Discharge Date (yyyy-mm-dd)`, `Discharge Status`)) %>%
  left_join(death) %>%
  mutate(death28d = case_when(
    `Discharge Status`=="DEATH" & 
      ymd(`Discharge Date (yyyy-mm-dd)`) - ymd(`Admission Date (yyyy-mm-dd)`) <= 28 &
      !is.na(`Discharge Status`)~1, 
      # unknown discharge status, unknown death28d status -> survival
      is.na(`Discharge Status`)~0,
      TRUE~0
    ),
     daystillevent = ymd(`Discharge Date (yyyy-mm-dd)`) - ymd(`Admission Date (yyyy-mm-dd)`),
     # right-censor
     daystillevent = ifelse(daystillevent > 28 & death28d == 0, 28, daystillevent),
     death28d = ifelse(daystillevent > 28 & death28d == 1, 0, death28d), 
     # replace na, ie discharged, with 28
     daystillevent = ifelse(is.na(daystillevent), 28, daystillevent),
     resp_death = ifelse(grepl("^J|^U071",`Death Cause (Main Cause)`) & death28d == 1, 1, 0)
    )

# ip.death %>% count(`Reference Key`) %>% count(n)

ip.death = ip.death %>%
  select(`Reference Key`, death28d, daystillevent, resp_death)

# colnames(ip_death_los)
writexl::write_xlsx(ip.death, './ip.death.xlsx')
```

# SDI

```{r}

ip.sdi = ip.nodrug %>%
  select(`Reference Key`, district = `District of Residence (system code)`) %>%
  group_by(`Reference Key`) %>%
  slice(which(district == first(district))) %>%
  ungroup() %>%
  distinct() %>%
  left_join(rename(sdi, district = Districtcode)) %>%
  select(-average_SDI_2)

writexl::write_xlsx(ip.sdi, path = "./ip.sdi.xlsx")
```


```{r}
ip.iptw.holder = ip.id %>%
  left_join(ip.sexage) %>%
  filter(!is.na(Sex) | !is.na(age)) %>%
  left_join(ip.sdi) %>%
  filter(!is.na(sdi_q))

ip.iptw.holder %>% nrow()

ip.iptw.holder %>% 
  left_join(ip.paxlovid) %>% 
  mutate(paxlovidlong = ifelse(!is.na(daystillpaxlovid), 1, 0),
         paxlovidshort = ifelse(daystillpaxlovid <= 4, 1, 0)) %>%
  
  left_join(ip.molnupiravir) %>%
  mutate(molnupiravirlong = ifelse(!is.na(daystillmolnupiravir), 1, 0),
         molnupiravirshort = ifelse(daystillmolnupiravir <= 4, 1, 0)) %>%
  filter(molnupiravirlong == 1) %>%
  # count(paxlovidlong) %>%
  filter(paxlovidlong != 1) %>%
  count(molnupiravirshort)
  # nrow()
  


ip.id %>% 
  left_join(ip.paxlovid) %>% 
  filter(is.na(daystillpaxlovid) | !(daystillpaxlovid > 4)) %>%
  mutate(paxlovid = ifelse(daystillpaxlovid <= 4, 1, 0)) %>%
  left_join(ip.molnupiravir) %>%
  filter(is.na(daystillmolnupiravir) | !(daystillmolnupiravir > 4)) %>%
  mutate(molnupiravir = ifelse(daystillmolnupiravir <= 4, 1, 0)) %>%
  
  # mutate(across(c("paxlovid", "molnupiravir"), ~replace_na(., 0))) %>%
  count(paxlovid, molnupiravir)
```


