---
title: "R Notebook"
output: html_notebook
---

# import libraries

```{r}
library(survival)
library(broom)
library(ipw)
library(survey)
library(comorbidity)
library(WeightIt)
library(cobalt)
library(tableone)
```


# 2022-04-26: episode death info is missing from hosp files; 
# please refer to data_import.Rmd for IP data import 
# check drug names (optional)

```{r}
unique(ip$`Drug Name`)[grepl("DEXAMETHASONE", unique(ip$`Drug Name`))]
```

# dissociate drug use and ip data (unique by HN number)

```{r}
# specify icd9 code column index
diag_cols = 35:49

df_ip = ip %>%
  select(-`Drug Name`) %>%
  distinct() %>%
  # exclude pregnancy
  mutate(pregnancy = ifelse(if_any(diag_cols, ~between(., "630", "679.99")), T, F),
         pregnancy = ifelse(is.na(pregnancy), F, pregnancy)) %>%
  filter(!pregnancy) %>%
  select(-pregnancy)
```

# identification dataframe of IP patients' first admissions only

```{r}
ip_first_by_patient_id = df_ip %>%
  select(`Reference Key`, `HN Number`, `Admission Date (yyyy-mm-dd)`, `Discharge Date (yyyy-mm-dd)`) %>%
  group_by(`Reference Key`) %>%
  slice(which.min(ymd(`Admission Date (yyyy-mm-dd)`))) %>%
  ungroup()
```

# medication status of IP epsiodes with IP data only

```{r}
ip_drug_use = ip %>%
  # exclude pregnancy
  mutate(pregnancy = ifelse(if_any(diag_cols, ~between(., "630", "679.99")), T, F),
         pregnancy = ifelse(is.na(pregnancy), F, pregnancy)) %>%
  filter(!pregnancy) %>%
  select(-pregnancy) %>% 
  select(`HN Number`, `Drug Name`) 

medIPdf = 
  ip_drug_use %>%
  impute_meds() %>%
  # baricitinib
  impute_meds_severe_covid() %>%
  select(-`Drug Name`) %>%
  distinct()

# aggregate by HN Number, get 1 if possible
medIPdf = aggregate(medIPdf, by = list(medIPdf[, "HN Number"]), FUN = "max") %>%
  select(-Group.1)

# severe covid-related meds status
# dexamethasone
dexamethasone_shortlist = steroid %>%
  select(`Reference Key`, `Prescription Start Date`, `Drug Name`, `Route`) %>%
  filter(grepl("DEXAMETHASONE", `Drug Name`) & (`Route` %in% c("ORAL","PARENTERAL"))) %>%
  left_join(ip_first_by_patient_id,.) %>%
  filter(ymd(`Prescription Start Date`) >= ymd(`Admission Date (yyyy-mm-dd)`) & 
           ymd(`Prescription Start Date`) <= ymd(`Discharge Date (yyyy-mm-dd)`)) %>%
  select(`Reference Key`) %>%
  distinct()

# combine with OP meds status (incomplete)
# 1. match by reference key &
# 2. appointment date <= (first) admission date
medOPdf_shortlist = medOPdf %>%
  left_join(ip_first_by_patient_id, by = "Reference Key") %>%
  filter(ymd(`Appointment Date (yyyy-mm-dd)`) <= ymd(`Admission Date (yyyy-mm-dd)`)) %>%
  select(c("Reference Key", "paxlovid", "aspirin", "ace_inhibitors", "beta_blockers", "calcium_channel_blockers", "statins", "molnupiravir"))

# combine with administration route info on dexamethasone form steroid
medIPdf_shortlist = medIPdf %>%
  left_join(ip_first_by_patient_id, by = "HN Number")

common_cols = intersect(colnames(medOPdf_shortlist), colnames(medIPdf_shortlist))
medIOPdf = bind_rows(medOPdf_shortlist %>% select(common_cols), 
                     medIPdf_shortlist %>% select(c(common_cols, "baricitinib")))

# aggregate medIOPdf by reference key
medIOPdf = aggregate(medIOPdf, by = list(medIOPdf[, "Reference Key"]), FUN = "max") %>%
  select(-Group.1) %>%
  # left-join with dexamethasone meds status 
  mutate(dexamethasone = ifelse(`Reference Key` %in% dexamethasone_shortlist$`Reference Key`, 1, 0)) %>%
  # as.factor binary cols
  mutate_at(2:length(colnames(medIOPdf)), factor) %>%
  mutate_at(2:length(colnames(medIOPdf)), ~replace_na(., 0))

```

# comorbs, change to comorbidity package

```{r}
# specify icd9 code column index
diag_cols = 35:49

ip_diag = df_ip %>% 
  select(`HN Number`, diag_cols) %>%
  impute_comorbs(., diag_cols - 33) 

ip_diag = aggregate(ip_diag, by = list(ip_diag[, "HN Number"]), FUN = "max") %>%
  select(-Group.1)

charlson_diag = df_ip %>% 
  select(`HN Number`, diag_cols) %>%
  # filter(`HN Number` %in% ip_first_by_patient_id$`HN Number`) %>%
  pivot_longer(cols = diag_cols - 33 , names_to = "diagnosis", values_to = "code") %>%
  drop_na() %>%
  group_by(`HN Number`) %>%
  distinct() %>%
  ungroup() %>%
  mutate(id = 1:n())

# CCI
charlson = comorbidity::comorbidity(x = as.data.frame(charlson_diag), 
                                   id = "id", 
                                   code = "code", 
                                   map = "charlson_icd9_quan", 
                                   assign0 = F) 
# charlson comorbs and cci by HN Number
charlson = charlson %>%
  left_join(select(charlson_diag, `HN Number`, id))
charlson = aggregate(charlson, by = list(charlson[, "HN Number"]), FUN = "max") %>%
  select(-Group.1)
charlson = charlson %>% 
  rowwise() %>% 
  mutate(cci = sum(c_across(mi:aids))) %>%
  mutate(cci = case_when(
    cci == 0~ "0",
    cci < 3~ "1-2",
    cci < 5~ "3-4",
    cci >= 5~ "5+",
    TRUE ~ "0"
  ))
cci = charlson %>%
  select(`HN Number`, cci)
# writexl::write_xlsx(charlson_diag, path = "./charlson_diag.xlsx")

# # charlson comorbidity index by HN Number
# cci = as.data.frame(list("cci" = comorbidity::score(charlson, weights = NULL, assign0 = FALSE)))
# cci = cci %>%
#   mutate(id = 1:n()) %>%
#   left_join(select(charlson_diag, `HN Number`, id),by = "id")
#   
# cci = aggregate(.~`HN Number`, cci, sum) %>%
  # mutate cci level
  # mutate(cci = as.factor(cci)) 
  
  
# write icd9 long format data to .xlsx
# writexl::write_xlsx(ip_diag, path = "./ip_icd9.xlsx")

ip_diag = ip_diag %>%
  left_join(cci, by = "HN Number") %>%
  # replace_na of ci
  mutate(cci = ifelse(is.na(cci), "0", cci))
  # select(-diagnosis, -id, -code)
  
unique(ip_diag$cci)

# check if there is any NA, missing data
ip_diag %>% filter_all((any_vars(is.na(.))))  

```



# IP data joins

# table 1 descriptive/SMD

```{r}
# hhttps://cran.r-project.org/web/packages/tableone/vignettes/smd.html#:~:text=The%20standardized%20(mean)%20difference%20is,and%20after%20propensity%20score%20matching.

ip_tab1_df = df_ip %>%
  mutate(`Admission Age (Year) (episode based)` = as.numeric(`Admission Age (Year) (episode based)`),
         ) %>%
  select(`Reference Key`, `HN Number`,`Admission Date (yyyy-mm-dd)`,
         `Admission Age (Year) (episode based)`, Sex) %>%
  # first admissions only
  filter(`HN Number` %in% ip_first_by_patient_id$`HN Number`) %>%
  # left_join ip_diag to df_ip by HN Number
  left_join(ip_diag, by = "HN Number") %>%
  # left join IOP meds status 
  left_join(medIOPdf, by = "Reference Key") %>%
  # impute age groups
  mutate(
    age = as.numeric(`Admission Age (Year) (episode based)`),
    age_group = 1,
    age_group = case_when( 
      age < 65 ~ "under65",
      (age >= 65 & age < 75) ~ "65-74",
      age >= 75~"over74",
      TRUE ~ as.character(age_group)
    )) %>%
  select(-`Admission Age (Year) (episode based)`)

# # optional, left join IP meds data only 
# ip_tab1_df = ip_tab1_df %>%
#   left_join(medIPdf_shortlist)

# # check if there is any NA, missing data
# ip_tab1_df %>% filter_all((any_vars(is.na(.))))  

# write excel by patient data
writexl::write_xlsx(ip_tab1_dfholder, path = "./ip_df.xlsx")
```



```{r}

ip_tab1_dfholder = ip_tab1_df %>%
  select(-`Reference Key`, -`HN Number`, -`Admission Date (yyyy-mm-dd)`) %>%
  # replace na of comorbs cols
  mutate_at(2:(length(colnames(.))-2), ~replace_na(., 0)) %>%
  # remove NA in sex and age
  drop_na() %>%
  # numeric binary to factor cateogorical cols
  mutate(across(!age, factor)) %>%
  select(-age)
  

# check if there is any NA, missing data
ip_tab1_dfholder %>% filter_all((any_vars(is.na(.))))  

ip_tab1_p = CreateTableOne(data = ip_tab1_dfholder, strata = "paxlovid", test = F)
print(ip_tab1_p, smd = T)
# ip_tab1_m = CreateTableOne(data = ip_tab1_dfholder, strata = "molnupiravir", smd = T)
# ip_tab1_m
```

# 28d mortality and length of stay

```{r}
# subselect relevant cols
ip_death_los = df_ip %>%
  select(
         # keys
         `Reference Key`, `HN Number`, 
         # death information, death if discharge status == death, discharge date becomes death date
         `Admission Date (yyyy-mm-dd)`, `Discharge Date (yyyy-mm-dd)`,  `Discharge Status`, 
         # length of stay LOS in days
         `Length of Stay`
         ) %>%
  filter(ymd(`Admission Date (yyyy-mm-dd)`) >= "2022-02-16" & ymd(`Admission Date (yyyy-mm-dd)`) <= "2022-03-31")  

  # left_join with table 1 data
  # %>% left_join(ip_first_by_patient_id, .)

# impute 28d death
ip_death_los = ip_death_los %>%
  mutate(death28d = case_when(
    `Discharge Status`=="DEATH" & 
      ymd(`Discharge Date (yyyy-mm-dd)`) - ymd(`Admission Date (yyyy-mm-dd)`) <= 28 &
      !is.na(`Discharge Status`)~1, 
      # unknown discharge status, unknown death28d status
      is.na(`Discharge Status`)~-1,
      TRUE~0
    ),
     daystilldeath = ymd(`Discharge Date (yyyy-mm-dd)`) - ymd(`Admission Date (yyyy-mm-dd)`),
     # right-censor
     daystilldeath = ifelse(daystilldeath > 28, 28, daystilldeath),
     # replace na with 0
     daystilldeath = ifelse(is.na(daystilldeath), 0, daystilldeath))
colnames(ip_death_los)
```

# propensity and IPTW by ipw/weightit

```{r}
# as.factor cateogorical cols and left_join ip_death_los data
ip_propensity = ip_tab1_df  %>%
  
  # left-join death data
  left_join(select(ip_death_los, `HN Number`, death28d)) %>%
  # discard unclear death status, ie == -1
  filter(death28d != -1) %>%
  mutate(across(2:length(colnames(ip_tab1_df)), factor)) %>%
  select(-`Reference Key`, -`HN Number`, -`Admission Date (yyyy-mm-dd)`) %>%
  # rename(age = `Admission Age (Year) (episode based)`) %>%
  # replace comorbs and meds stats cols NA with 0
  mutate_at(diabetes:dexamethasone, ~replace_na(., 0)) %>%
  # drop sex and age NA
  drop_na() %>%
  # impute age groups
  mutate(
    age = as.numeric(age),
    age_group = 1,
    age_group = case_when( 
      age < 65 ~ "under65",
      (age >= 65 & age < 75) ~ "65-74",
      age >= 75~"over74",
      TRUE ~ as.character(age_group)
    )) %>%
  select(-age)

# check if there is any NA, missing data
ip_propensity %>% filter_all((any_vars(is.na(.))))
# write to xlsx
writexl::write_xlsx(ip_propensity, path = "./ps_analysis_v4.xlsx")
```


```{r}
x_factors = colnames(ip_propensity)[! colnames(ip_propensity) %in% c("paxlovid", "death28d")]
  
# weightit 
function_call<-paste0("weight <- weightit(paxlovid ~ ",paste(x_factors, collapse = "+"), 
              ", data = ip_propensity, method = \"ps\", 
              estimand = \"ATT\")"
              )
eval(parse(text = function_call))

# diagnostics of weighting

bal.tab(weight, un = T)
summary(weight)



clus <- svydesign(id =~ 1, weights = weight$weights, data = ip_propensity)
res <- svyglm(death28d ~ paxlovid, design = clus,family = binomial)
ps_holder = tidy(res, exponentiate = T, conf.int = T, conf.level = .95)
ps_holder %>% filter(term != "(Intercept)")

writexl::write_xlsx(ps_anlysis, path = "./ps_analysis_ATT.xlsx")
```

# weighted descriptive stats

```{r}
ip_tab1_p_weighted = svyCreateTableOne(data = clus, strata = "paxlovid", test = F)
print(ip_tab1_p_weighted, smd = T)
```

# table 2b descriptive 

```{r}
# select only sex, age, obesity, diabetes, paxlovid and molnupiravir
ip_2b = ip_death_los %>% 
  select(daystilldeath, death28d, Sex, `Admission Age (Year) (episode based)`, paxlovid, molnupiravir) %>%
  # impute 65+
  mutate(over64 = ifelse(`Admission Age (Year) (episode based)` >= 65, 1 , 0)) %>%
  mutate(across(!daystilldeath, factor)) %>%
  select(-`Admission Age (Year) (episode based)`)

ip_tab2b = CreateTableOne(data = ip_2b, strata = c("paxlovid"), test = F)
print(ip_tab2b, smd = T)
```

# (univariate) hazard ratio per predictor

```{r}

ip_hazard = ip_death_los %>%
  select(4:30, 34:35) %>%
  # impute age groups
  mutate(
    age = as.numeric(`Admission Age (Year) (episode based)`),
    age_group = 1,
    age_group = case_when( 
      age < 65 ~ "under65",
      (age >= 65 & age < 75) ~ "65-74",
      age >= 75~"over74",
      TRUE ~ as.character(age_group)
    )) %>%
  select(-age, -`Admission Age (Year) (episode based)`)

ip_hazard %>% filter_all(any_vars(is.na(.)))


model = as.data.frame(c())
# iterate over cols
for(x in colnames(ip_hazard)){
  
  iformula <- as.formula(sprintf("Surv(daystilldeath, death28d) ~ %s", x))
  coxfit_paxlovid <- coxph(formula = iformula, 
                data = ip_hazard)
  model = bind_rows(model, tidy(coxfit_paxlovid, exponentiate = T, conf.int = T, conf.level = .95))
}
model %>% filter(p.value <= 0.05, conf.low > 1)
```

# multivariate model

```{r}
coxfit_paxlovid <- coxph(formula = Surv(daystilldeath, death28d) ~ Sex+ cci+age_group+paxlovid+molnupiravir, 
                data = ip_hazard)
tidy(coxfit_paxlovid, exponentiate = T, conf.int = T, conf.level = .95)
plot(survfit(coxfit_paxlovid))
```

# LOS

```{r}

ggplot(ip_death_los, aes(x = as.numeric(`Length of Stay`), fill = as.factor(paxlovid))) +
  geom_histogram()
```

